Clustering exploration for `r params$library_id`

### 2024-11-21 analyses/cell-type-ewings/template_notebooks/clustering-workflow/01-clustering-metrics.Rmd
remove extra cluster code

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  cluster_results_file: "SCPCL000822_clusters.tsv"
---

@@

Louvain clusters are calculated using the following range of parameters:

- Nearest neighbors: 5, 10, 15, 20, 25, 30, 35, and 40
- Resolution: 0.5, 1, 1.5

Leiden clusters are calcualted using the following range of parameters:

- Objective function: CPM and modularity 
- Nearest neighbors: 5, 10, 15, 20, 25, 30, 35, and 40
- Resolution: 0.5, 1, 1.5

- Metrics are then calculated to evaluate clustering results across all parameters: 
  - Silhouette width 
  - Cluster purity 
  - Cluster stability

@@

```{r}
# read in input sce files
sce <- readr::read_rds(params$sce_file)
```

## Calculate clusters

```{r}
# louvain clustering across all parameters 
louvain_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "louvain",
  nn = seq(5, 40, 5),
  resolution = c(0.5, 1, 1.5)
) |> 
  dplyr::bind_rows() |> 
  # add in objective function so we can combine with leiden results
  dplyr::mutate(objective_function = NA_character_)

# leiden cpm clustering
leiden_cpm_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "leiden",
  nn = seq(5, 40, 5),
  resolution = c(0.001, .005, .01),
  objective_function = "CPM"
) |> 
  dplyr::bind_rows()

# leiden modularity clustering
leiden_mod_results <- rOpenScPCA::sweep_clusters(
  sce,
  algorithm = "leiden",
  nn = seq(5, 40, 5),
  resolution = c(0.5, 1, 1.5),
  objective_function = "modularity"
) |> 
  dplyr::bind_rows()
```

```{r}
# join all clustering results 
all_cluster_results <- dplyr::bind_rows(list(louvain_results, leiden_cpm_results, leiden_mod_results)) |> 
  # add column for method + objective function 
  # use this for splitting plots throughout report
  dplyr::mutate(cluster_method = glue::glue(
    "{algorithm}-{objective_function}"
  ) |> 
    stringr::str_remove("-NA")
  )
```

@@

```{r}
# get umap embeddings and combine into a data frame with cluster assignments
umap_df <- sce |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  # replace UMAP.1 with UMAP1 and get rid of excess columns
  dplyr::select("cell_id" = barcodes, UMAP1 = UMAP.1, UMAP2 = UMAP.2) |>
  dplyr::left_join(all_cluster_results, by = "cell_id")

@@

## Export clusters

```{r}
# export all clustering results
readr::write_tsv(all_cluster_results, params$cluster_results_file)
```

>>>>

---
title: "Clustering exploration for `r params$library_id`"
author: Ally Hawkins
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: "hide"
params:
  library_id: "SCPCL000822"
  sce_file: "../../../../data/current/SCPCP000015/SCPCS000490/SCPCL000822_processed.rds"
  cluster_results_file: "../../results/clustering/SCPCS000490/SCPCL000822_cluster-results.tsv"
---

@@

The following metrics are calculated and shown across all parameters: 
  - Silhouette width 
  - Cluster purity 
  - Cluster stability

@@

```{r}
# read in input sce files
sce <- readr::read_rds(params$sce_file)

# read in results file 
cluster_results_df <- readr::read_tsv(params$cluster_results_file)
```

## Clustering summary

```{r, results='asis'}
cluster_results_df |> 
  split(cluster_results_df$cluster_method) |> 
  purrr::imap(\(df, method){
    res_print <- unique(df$resolution) |> 
      paste0(collapse = ", ")
    nn_print <- unique(df$nn) |> 
      paste0(collapse = ", ")
    glue::glue("
               {method} clustering was performed with the following parameters:

                - Nearest neighbors: {nn_print}
                - Resolution: {res_print}"
               )
  })
```

@@

```{r}
# get umap embeddings and combine into a data frame with cluster assignments
umap_df <- sce |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  # replace UMAP.1 with UMAP1 and get rid of excess columns
  dplyr::select("cell_id" = barcodes, UMAP1 = UMAP.1, UMAP2 = UMAP.2) |>
  dplyr::left_join(cluster_results_df, by = "cell_id")

@@
